---
title: "Homework 9"
author: "Harshetha Haritharan - hh29685"
output: 
  pdf_document:
    latex_engine: xelatex
date: "2025-04-19"
---
GitHub Repository: <https://github.com/harshetha-haritharan/HW9>

## Problem 1
```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
solder <- read.csv("solder.csv")
```

## Part A
```{r, echo=FALSE}
# Boxplots showing relationships

ggplot(solder, aes(x = Opening, y = skips)) +
  geom_boxplot() +
  labs(title = "Relation of Opening Size to Solder Skips",
       x = "Opening Size", y = "Number of Skips")

ggplot(solder, aes(x = Solder, y = skips)) +
  geom_boxplot() +
  labs(title = "Relation of Thickness of Alloy to Solder Skips",
       x = "Thickness of Alloy", y = "Number of Skips")
```
The Relation of Opening Size to Solder Skips boxplot shows that smaller opening size on the solder gun have an association with a higher number of skips while large openings have an association with fewer skips. From this, it can be suggested that the various opening size on the solder gun is related to the number of skips.

This Relation of Thickness of Alloy to Solder Skips boxplot shows that thinner solder has an association with more skips when being compared to thick solder, therefore suggesting that that solder thickness is related to the number of skips.

## Part B
```{r regression-table, message=FALSE, warning=FALSE, echo = FALSE}
library(moderndive)
# Fit model
model <- lm(skips ~ Opening * Solder, data = solder)

# Create regression table 
get_regression_table(model)
```
## Part C

The baseline number of skips when Opening = L and Solder = Thick is 0.393, meaning that the model predicts that a board manufactured with a large opening and thick solder will have 0.393 skips on average. 

When Opening = M and Solder = Thick, the average number of skips increases by 2.407, leading to a total of 2.800 skips.

When Opening = S and Solder = Thick, the average number of skips increases by 5.127, leading to a total of 5.520 skips.

When Opening = L and Solder = Thin, the average number of skips increases by 2.280, leading to a total of 2.673 skips.

When Opening = M and Solder = Thin, the number of skips is 0.740 (interaction coefficient) fewer that what we would expect if we just added the two individual effects. The effect of the thin solder is not as severe when used with a medium opening compared to when it is with a large opening. 

When Opening = S and Solder = Thin, the number of skips is 9.653 (interaction coefficient) greater that what we would expect if we just added the two individual effects. The effect of thin solder is more severe when used with a small opening compared to when it is used with a large opening. 

## Part D
If I had to recommend a combination of Opening size and Solder thickness to AT&T based on this analysis I would recommend a large opening size with a thick solder. This is the baseline group value in the regression model and its value, 0.393, has the lowest predicted number of skips. When compared to the other combination, the others have higher skip amounts, notably those with small opening or thin solder. Small opening or thin solder both independently have a high number of skips and when used together they have the highest number of skips. Therefore, large opening and thick solder is the best combination to minimize the number of skips in the manufacturing process.


## Problem 2
## Part A
```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
groceries <- read_csv("groceries.csv", show_col_types = FALSE)

# Calculate average price
avg_price <- groceries %>%
  group_by(Store) %>%
  summarize(avg_price = mean(Price, na.rm = TRUE)) %>%
  arrange(desc(avg_price))

# Create plot
ggplot(avg_price, aes(x = reorder(Store, avg_price), y = avg_price)) +
  geom_col(fill = "red") +
  coord_flip() +
  labs(
    title = "Average Product Price by Store",
    x = "Store",
    y = "Average Price in Dollars")
```
This bar plot shows the average price in dollars of all of the products sold at each store. It can be seen that there is variation across the stores and specific stores such as Whole Foods and Wheatsville Food Co-op have notably higher average prices compared to the other stores. 

## Part B
```{r, echo=FALSE}
# Count how many stores are selling each product
product_counts <- groceries %>%
  group_by(Product) %>%
  summarize(n_stores = n()) %>%
  arrange(desc(n_stores))

# Create plot
ggplot(product_counts, aes(x = n_stores, y = reorder(Product, n_stores))) +
  geom_col(fill = "orange") +
  labs(
    title = "Number of Stores Carrying Each Product",
    x = "Number of Stores",
    y = "Product"
  ) 
```
This bar plot shows that number of stores that carry of the products listed. It can be seen that common goods such as milk and eggs are in more stores, compared to more unique items which are in a lesser number of stores. This can help us understand why the average store price comparisons seen in the Part A plot may be not be a good representation as not all the stores sell all of the products. 

## Part C
```{r, echo=FALSE}
# Fit model
model_type <- lm(Price ~ Product + Type, data = groceries)

# Find summary and confidence intervals
coefs <- summary(model_type)$coefficients
ci <- confint(model_type)

type_rows <- grep("Type", rownames(coefs))

# Combine estimate and confidence interval
table <- data.frame(
  Term = rownames(coefs)[type_rows],
  Estimate = round(coefs[type_rows, 1], 2),
  Lower_95CI = round(ci[type_rows, 1], 2),
  Upper_95CI = round(ci[type_rows, 2], 2)
)

print(table, row.names = FALSE)


```
Compared with ordinary grocery stores (like Albertsons, HEB, or Krogers), convenience stores charge somewhere between 0.41 and $0.92 dollars more for the same product.

## Part D
```{r, echo=FALSE}
model_store <- lm(Price ~ Product + Store, data = groceries)

# Get summary and confidence intervals
coefs <- summary(model_store)$coefficients
ci <- confint(model_store)

# Get rows related to Store
store_rows <- grep("^Store", rownames(coefs))

# Create table
store_table <- data.frame(
  Store = rownames(coefs)[store_rows],
  Estimate = round(coefs[store_rows, 1], 2),
  Lower_95CI = round(ci[store_rows, 1], 2),
  Upper_95CI = round(ci[store_rows, 2], 2)
)

# Sort by estimated price difference
store_table <- store_table[order(store_table$Estimate), ]
print(store_table, row.names = FALSE)
```
One of the stores with the lowest estimated prices are Walmart with prices that average $0.99 less than the baseline store. Walmart has a 95% confidence interval of [-$1.45, -$0.53] meaning it is statistically significant. The other store is Kroger Fresh Fare with prices that are $0.90 less and it has a 95% confidence interval of [-$1.36, -$0.44] which is also statistically significant. Next, one of the stores with the highest estimated prices are Whole Foods as it averages $0.36 more than the baseline and has a confidence interval of [$0.02, $0.71] meaning it is statistically significant. The other store is Wheatsville food Co-op with prices $0.29 higher than the average with a confidence interval of [-$0.06, $0.64] and since it includes zero, it is not statistically significant. 

## Part E
```{r, echo=FALSE}
# Get H-E-B and Central Market values
heb_est <- coefs["StoreH-E-B", 1]
heb_ci <- ci["StoreH-E-B", ]
cm_est <- coefs["StoreCentral Market", 1]
cm_ci <- ci["StoreCentral Market", ]

# Compute difference 
difference <- cm_est - heb_est
cat("H-E-B estimate:", round(heb_est, 2), 
    "with 95% CI [", round(heb_ci[1], 2), ",", round(heb_ci[2], 2), "]\n")

cat("Central Market estimate:", round(cm_est, 2), 
    "with 95% CI [", round(cm_ci[1], 2), ",", round(cm_ci[2], 2), "]\n")

cat("Difference between Central Market and H-E-B:", round(difference, 2), "\n")

# Compare range for all stores
store_rows <- grep("^Store", rownames(coefs))
store_estimates <- coefs[store_rows, 1]
spread <- max(store_estimates) - min(store_estimates)

cat("Total price spread across stores:", round(spread, 2), "\n")
```
From the model, it can be seen that H-E-B has a coefficient of -0.65 (95% Cl: [-0.95, -0.35]) and Central Market has a coefficient of -0.57 (95% cl: [-0.92, -0.23]). Therefore the estimated price difference between Central Market and H-E-B is 0.07, meaning that Central Market charges approximately $0.08 more than H-E-B for the same product. This supports the possibility that Central Market charges a similar amount to H-E-B for the same product. When compared to other store coefficients such as Walmart (-$0.99) to CVS ($0.19),the $0.08 difference seems minor. Furthermore, the Central Market and H-E-B gap is not major as the total price spread across the stores is calculated to be $1.36, which is more than $1.00.

## Part F
```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(effectsize)

groceries <- groceries %>%
  mutate(Income10K = Income / 10000)

model_income <- lm(Price ~ Product + Income10K, data = groceries)
coefs_income <- summary(model_income)$coefficients["Income10K", ]
coefs_income
standardize_parameters(model_income) %>%
  filter(Parameter == "Income10K")
```
Consumers in poorer ZIP codes may pay slightly more for the same product, as indicated by the negative coefficient for Income10K (–0.014). This means that for every $10,000 increase in ZIP code income, the average product price decreases by about 1.4 cents. However, the p-value is 0.144, which is not statistically significant at the 5% level, so we cannot confidently conclude that this relationship holds in the broader population. When using standardized income (z.Income10K), the standardized coefficient is –0.03, with a 95% confidence interval of [–0.07, 0.01]. A one-standard deviation increase in ZIP code income is associated with a 0.03 standard deviation decrease in product price, but again, the relationship is not statistically significant.


## Problem 3

## Statement A: ZIP codes with a higher percentage of minority residents tend to have more FAIR policies per 100 housing units.
This statement is true because Figure A1 shows that there is a positive and linear trend among the percentage of minority residents and the amount of FAIR policies. Using values from regression model_A, it can be seen that the coefficient for minority residents is 0.014, meaning that for every 1% point increase in the percentage of minority residents in each ZIP code, the amount of FAIR plans per 100 housing units increases by 0.014 on average. The p-value is also less than 0.001 meaning that it is very significant. The R-squared value is 0.516, which means that 51.6% of the variation in FAIR policy rates across ZIP codes can be explained by the percentage of minority residents. All of this evidence supports that there is a positive association, making this a true statement. 

## Statement B: The evidence suggests an interaction effect between minority percentage and the age of the housing stock in the way that these two variables are related to the number of FAIR policies in a ZIP code.
Statement B is undecidable because while the statement claims there is an interaction effect between minority percentage and housing age in predicting FAIR policies, no model or figure directly tests or shows this. Model B only examines whether age is related to minority percentage, and the association is weak because R² = 0.06 and p = 0.125. Figure B1 supports this, showing no clear trend as the points are spread out and the fitted line is flat. However, no model includes an interaction term between minority and age, and no figure shows how FAIR policy rates vary across combinations of those two variables. To correct it we can say that we do not currently have evidence for an interaction between minority percentage and housing age in predicting FAIR policies. To support this claim, it would be good to make a regression model with a minority * age interaction term and a plot showing how the relationship between minority percentage and FAIR policy uptake changes across different housing ages.

## Statement C: The relationship between minority percentage and number of FAIR policies per 100 housing units is stronger in high-fire-risk ZIP codes than in low-fire-risk ZIP codes.
This statement is undecidable because while Figure C1 appears to show a slightly steeper slope for high-fire-risk ZIP codes, the difference is small, and the lines are pretty similar. More importantly, in Model C, the interaction term for minority and low fire risk has a coefficient of –0.001 with a 95% confidence interval of [–0.012, 0.010]. Because this confidence interval includes 0, we cannot conclude that there is a statistically significant difference in the minority effect between high and low fire risk ZIPs. Therefore, there is not enough statistical evidence to claim that the relationship between minority percentage and FAIR policy rates differs across fire risk levels and to undertstand the situation better, we would need stronger evidence that the interaction term is significantly different from zero.

## Statement D: Even without controlling for any other variables, income “explains away” all the association between minority percentage and FAIR policy uptake.
This statement is false because although income does reduce the strength of the association between minority percentage and FAIR policy uptake, it does not explain it entirely. In model D1, where there is no control for any other variables, the coefficient for minority is 0.014, with a 95% confidence interval of [0.006, 0.023] and is statistically significant. After controlling for income in model D2, the coefficient for minority decreases to 0.010, and is still statistically significant because the 95% confidence interval [0.004, 0.017] excludes 0. Even though income is also statistically significant because p = 0.041, the inclusion of income does not eliminate the effect of the minority variable. The statement would be more accurate if it was worded as Controlling for income reduces the association between minority percentage and FAIR policy uptake, but does not eliminate it. Minority percentage continues to show an independent effect. This is supported by model D2, where the minority coefficient remains significant even after controlling for income. Its value of 0.010 shows that the positive connection with FAIR policy uptake still exists. The decrease from 0.014 in model D1 to 0.010 in model D2 suggests that income explains part of the relationship, but not all of it.

## Statement E: Minority percentage and number of FAIR policies are still associated at the ZIP code level, even after controlling for income, fire risk, and housing age.
This statement is true because even after being controlled for multiple factors, the minority percentage is still significantly associated with FAIR policies at the ZIP code level. In model E, which is the the fully adjusted model, the coefficient for minority is 0.008, and its 95% confidence interval [0.002, 0.015] excludes 0, showing that there is a statistically significant positive relationship. The model includes controls for income, fire risk, and housing age, but only the minority and fire risk values are statistically significant. The R-squared value is 0.662, showing that the model explains a large portion of the variance in FAIR policy rates. This gives strong evidence that the association between minority percentage and FAIR policy use is not explained by other ZIP-level characteristics such as fire risk or income.